---
title: "brainstorming"
output: pdf_document
date: "2025-06-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Debugging:

-   Task 1 issues:

    -   I am unsure why [Projection ...] is printing to console.

    -   I am unsure why keys are printing

-   Task 2 issues:

    -   Merge returns a 2 column data frame, but I only want one column per species.

        -   solution: Remove duplicate rows that have no species for row 1

    -   

## Preparing genes for analysis

```{r}
library(stringr)
library(readr)
library(dplyr)

#geneCollector accesses and cleans loss_summ_data for each species to be examined. Then, it  
geneCollector <- function(){
 # Access losssum.tsv
    setwd("~/meyerLab/psuedogeneCollector/data")
      files <- list.files(
        path =".", 
        recursive=TRUE, 
        pattern = "loss_summ_data",
        full.names=TRUE)
        
    species_gene <- grabData(files)
    genes <- updateUniqueGenes(files)
    
    # ###Testing: 
    # cat("\n\n\n Testing species_gene dictionary: \n") # --> Test passed
    # for (key in names(species_gene)) {
    #   value <- species_gene [[key]]
    #   cat( "\n\n\n\n\ ","Key:", key, " Value: ", " \n\n\n\n")
    #   print(value[1:10,]) 
    #   }
    #   
    # cat("\n\n\n Testing for genes: \n") # --> Test passed 
    # print(genes[1:10])
    # head(species_gene[[1]]) #looking at first few rows of the df of key 1
    
    return(list(species_gene = species_gene, genes = genes))
}

#' Obtains and cleans loss_summ_data
#' @param files --> Vector containing absolute paths to every instance of loss_summ_data
#' @return A list formatted as a dictionary --> key = character (species name), value = df (clean loss_summ_data)
#' @author Tyler gruver
    grabData <- function(files){
      species_gene <- list() #key == species name, value == .tsv file containing losssum data
    #  counter <- 0
      
    
      
      
      for(i in 1:length(files)){
        temp <- str_locate_all(files[i], "/")[[1]] #temp returns a list of matrixes. [[1]] grabs the matrix  at position 1... temp will only ever be a list of length 1 

        # counter <- counter + 1
        # if(counter > 5) break   
        # 
          
        #Setting key-value pairs
        name <- substring(files[i], temp[3, 1] +1 , temp[4, 1] -1)
        key <- paste("Key: ", name , sep = " ")
        value <- cleanData(files[i]) #returns clean .tsv
        species_gene[[key]] <- value
       
        #cat("\n\n\n\n ", counter, " iterations /n/n/n/n/n")
        
      }  
      
    
        return(species_gene)
    }
    
#' cleans loss_summ_data
#' @param f --> Vector containing absolute paths to a single instance of loss_summ_data
#' @return A df --> Clean loss_summ_data
#' @author Tyler gruver
cleanData <- function(f){
   df <- read_tsv(f, show_col_types = FALSE)
  clean_df <- df[df$PROJECTION == "GENE",]
  return(clean_df[, -1])
}

#' Creates a list of unique genes seen under every instance of loss_summ_data
#' @param files --> Vector containing absolute paths to a single instance of loss_summ_data 
#' @return A list containing every unique gene under every instance of loss_summ_data
#' @author Tyler gruver    
  updateUniqueGenes <- function(files){
    genes <- character(length=0)
    
#    counter = 0

    
    for(f in files) {
      #counter <- counter + 1
      
      df <- read_tsv(f, show_col_types = FALSE)
      genes <- union(genes, df[,1])  # performs set union operation on genes character vector AND df[,1] column subset.
      
 #     if(counter > 5) break
      
        }
  
    return(genes)
  }

  
  
  var1 <- geneCollector()
    
```

# Task 2:

Create a new .tsv file.\
Add a col for every species

Add a row for every gene in genes

For every species, check for matches between every gene in their data frame for a match in genes. If a match is found, place element in df[species name, gene found]

```{r}
library(stringr)
library(readr)

#geneCollector accesses and cleans loss_summ_data for each species to be examined. Then, it  
geneCollector <- function(){
 # Access losssum.tsv
    setwd("~/meyerLab/psuedogeneCollector/data")
      files <- list.files(
        path =".", 
        recursive=TRUE, 
        pattern = "loss_summ_data",
        full.names=TRUE)
                
    species_gene <- grabData(files)
    genes <- updateUniqueGenes(files)
    
    mergeData(species_gene, genes) #generates tsv
    
    # ###Testing: 
    # cat("\n\n\n Testing species_gene dictionary: \n") # --> Test passed
    # for (key in names(species_gene)) {
    #   value <- species_gene [[key]]
    #   cat( "\n\n\n\n\ ","Key:", key, " Value: ", " \n\n\n\n")
    #   print(value[1:10,]) 
    #   }
    #   
    # cat("\n\n\n Testing for genes: \n") # --> Test passed 
    # print(genes[1:10])
    # head(species_gene[[1]]) #looking at first few rows of the df of key 1
    
    return(list(species_gene = species_gene, genes = genes))
}

  #Use left_join --> all data from left df will stay and data in the right df
  mergeData <- function(species_gene, genes) {
    write_tsv(file = "psuedogeneData_Merged")
    
  }
  
  
  geneCollector()

```

-   COL 2 should contain all gene names. Look through every gene name. For every new gene, append that gene to the list that stores genes. 
